### 2013-11-19 manuscripts/reply/reply-main.tex
Working on reply.
Moving error section to SI, and working on Intro.

Recently, the journal \emph{Evolution} has served as a medium for a discussion
of the potential pitfalls of approximate Bayesian methods of comparative
phylogeographical model choice and prior selection.
The discussion has centered around the approximate Bayesian model choice method,
\msb, which estimates the temporal distribution of divergences among
co-distributed pairs of taxa \citep{Huang2011} by sampling over all
unordered models of divergence.
\citet{Oaks2012} published results of simulation-based power analyses
showing that \msb can often be biased toward inferring models of temporally
clustered divergence times across taxon pairs.
\citet{Hickerson2013} published a response to \citet{Oaks2012} where they
present an empirical Bayes model-averaging approach that they conclude
circumvents the poor behavior of the method revealed by \citet{Oaks2012}.
These papers are largely in agreement, with \citet{Hickerson2013} reiterating
much of the discussion of \citet{Oaks2012} regarding the impact of broad
uniform priors on Bayesian model choice.
The main difference in perspectives are centered around
(1) the mechanism by which broad uniform priors cause the poor behavior of
\msb, and
(2) how to potentially mitigate this issue.

\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
is likely the low marginal likelihoods of parameter-rich models integrated over
vast parameter space with low probability of producing the data, yet relatively
high prior density \citep{Lindley1957}.
Note, this suggests the bias is extrinsic to \msb, and the numerical
approximation machinery of \msb could be sound.
\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
method's rejection algorithm is inefficient and will be increasingly biased as
the overall space of the model increases, either as a function of the number of
taxon pairs or the width of the uniform priors on nuisance parameters.
\citet{Hickerson2013} support their position by giving a probabilistic argument
that focuses on only one of the three prior models used by \citet{Oaks2012}.
We show that this argument is based on an empirical Bayesian reasoning and
questionable assumptions, and does not explain the bias of the method found in
many of the analyses of \citet{Oaks2012}. 
Furthermore, we present results of additional analyses that, when combined with
the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
models with more parameter space is a primary mechanism by which uniform priors
are causing the bias of \msb.

%This paragraph seems too aggressive. It also seems to repeat some of the previous points. I removed it for now.

To mitigate the bias, \citet{Oaks2012} suggest alternative prior probability
distributions on divergence models and nuisance parameters could increase the
likelihoods of models with more divergence time parameters.
They also describe simulation-based procedures for determining power, accuracy,
and robustness of the method given a dataset, and recommend any application of
\msb should be accompanied by such procedures.
Alternatively, \citet{Hickerson2013} present an approximate Bayesian
model-averaging approach for accommodating uncertainty in selecting among
empirically guided priors, and champion the method as a means of avoiding the
pitfalls raised by \citet{Oaks2012}.
We discuss errors in this approach that render some of their results invalid
and leave the remaining results difficult to interpret.
Furthermore, we follow the recommendations of \citet{Oaks2012} and present
simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
approach.
Our results demonstrate that the empirical Bayesian approach of
\citet{Hickerson2013} is biased toward clustered divergence models and
predisposed to excluding true parameter space.
The model-averaging approach provides another means by which the model can
``escape'' large parameter space, and the method's bias toward smaller models
remains.
This manifestation of the bias often excludes the true parameter space due to
the use of narrow uniform priors.
Furthermore we discuss the potential theoretical and practical problems of
empirical Bayesian model choice.
% which is used and encouraged by
% \citet{Hickerson2013} without discussing its potential complications.

%CWL:
%It may be important to point out the error made in their paper, but it can
%also be nit-picky, it depends on how the error impacts the overall message of
%our response.  We should strive to keep the response short and directed at the
%big picture items. If there is a way to still point out the error, then maybe
%some of this section should be used
\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
model-averaging approach, where they placed a discrete uniform prior over eight
different prior models (see Table 1 of \citet{Hickerson2013}).
However, there is a fundamental error in the methodology of
\citet{Hickerson2013}: their model mixes different time units.

Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
has one of two priors on the mean size of the descendant populations of each
taxon pair:
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
As described in \citet{Oaks2012}, the divergence time parameters of the model
implemented in \msb are scaled relative to a constant reference population
size, \myTheta{C}.
This reference population size is defined in terms of the upper limit of the
uniform prior on the mean size of the descendant populations,
\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
Thus, the model used by \citet{Hickerson2013} mixes two different units of
time.
In other words, some of their prior and posterior samples are in units of
$0.05/\mutationRate$ generations, whereas others are in units of
$0.02/\mutationRate$ generations.

The fact that their posterior samples are in different units makes the results
of \citet{Hickerson2013} difficult to interpret, and renders their
regression-adjusted results invalid.
A fundamental assumption of regression is that all of the values of the
response variable are in the same units.
Thus, the results in sections ``Using ABC Model Comparison to Weight
Alternative Priors for the Philippine Vertebrate Data'' and ``Improved Sampling
Efficiency by Prior Weighting Supports Asynchronous and Recent Divergence for
the Philippines Vertebrate Data'' and presented in Figure 2 of
\citet{Hickerson2013} should be disregarded.
The error is easily illustrated by re-plotting their results with the different
time units indicated (Figure~\ref{figJointPosteriorHickerson}).

@@

\section*{The potential implications of empirical Bayesian model choice}
\citet{Hickerson2013} repeatedly refer to the prior distributions used by
\citet{Oaks2012} as ``poorly selected.''
However, this is misleading, as \citet{Oaks2012} discuss in detail how they
selected their priors to reflect the large amount of uncertainty about the
model's parameters (see their section ``Specifying and simulating the joint
prior'').
\citet{Oaks2012} further discuss how the use of uniform prior distributions in
\msb for many of the model's parameters necessitates awkward broad priors to
avoid excluding the truth, \textit{a priori}.
What \citet{Hickerson2013} mean by ``poorly selected'' is that
\citet{Oaks2012} did not use their data to inform their initial prior
distributions (i.e., they did not use an empirical Bayesian approach).
\citet{Oaks2012} did use empirically informed priors when assessing both the
prior sensitivity of the method and determining the power of \msb under
real-world situations when a large amount of prior information is available. 
\citet{Oaks2012} discuss
the potential dangers of taking an empirical Bayesian approach to model choice
(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
and practical implications of using empirically informed priors for Bayesian
model choice.

@@

Whereas empirical Bayes approaches can provide powerful methods for parameter
estimation, a theoretical justification for empirical Bayes approaches to model
choice is questionable.
In Bayesian model choice, the primary goal is not to estimate parameters, but
to estimate the relative probabilities of candidate models.
In a simple example of Bayesian model choice where two candidate models, $M_1$
and $M_2$, are being compared, the goal is to estimate the posterior
probabilities of these two models.
Again, we can use Bayes' rule to calculate this as
\begin{equation}
    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
    M_2)p(M_2)}.
    \label{eq:bayesmodelchoice}
\end{equation}
By comparing Equations \ref{eq:bayesrule} and \ref{eq:bayesmodelchoice}, we
can see the fundamental differences between Bayesian parameter estimation and
model choice.
In Equation \ref{eq:bayesrule}, we can see that the posterior density of any
state $\myTheta{}$ of the model, is the prior density updated by the
probabiliity of the data given the state $\myTheta{}$ (the likelihood of
$\myTheta$).
The integral over the entire parameter space of the model, which is defined by
the priors, only appears as a normalizing constant in the denominator.
Thus, as long as the joint prior distribution contains the values of $\myTheta$
that maximize the probability of the data (i.e., the maximum likelihood) and the
data are strongly informative relative to the prior, the modes of the parameter
estimates will be relatively robust to prior choice, even if the shape of the
posterior is incorrect due to using the data to inform the priors.
This is why empirical Bayes can work well for estimating the values of model
parameters.

However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model
the updates the prior probability of that model to yield posterior
probability.
Thus the integral over the entire parameter space of the likelihood weighted by
the prior probability density is no longer a normalizing constant, but rather
is what informs the posterior probability of that model from the data.
As a result, Bayesian model choice tends to be much more senstitive to priors
than Bayesian parameter estimation.
Another important difference of Bayesian model choice that is illustrated by
Equations~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a parameter of the model, but rather a
single value that is a summary of the entire posterior distribution.
This further complicates the meaning of model posterior probabilities when
Bayes' rule is violated by empirically informing model priors.
Model posterior probabilities are inherently measures of our belief in
the models after our prior beliefs are updated by the data being analyzed.
Unlike model parameters, there are no true values for model probabilities,
and there are no measures of uncertainty in the posterior estimates. 
For empirical Bayesian parameter estimation, the hope is that by giving the
data more weight relative to the prior, the posterior distribution will be
peaked near the true value(s) of the model's parameter(s).
There is no such justification for empirical Bayesian model choice, because
there are no true model probabilities that we are trying to estimate.
Rather using the data twice will simply mislead our posterior
beliefs in the models being compared.

@@

\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
The main parameter of the \msb model is the vector of divergence
times for each of the taxon pairs,
$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
\citep{Oaks2012}.
\citet{Hickerson2013} argue that the dispersion index of this vector,
\vmratio{}, is a better model-choice estimator than the number of 
divergence time parameters within the vector,
\numt{}.
They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
This plot shows, not surprisingly, that \vmratio{} has very little information
about the number of divergences among taxa.
Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
biogeographically relevant than \numt{}.
% We struggle to follow this logic.
Certainly the maximum information is contained within the divergence time
vector that \vmratio{} is summarizing.
Clearly, the number of divergence time parameters within the vector and their
values is more informative than its variance (i.e., the dispersion index is not
a sufficient statistic for \divtvector).

\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
better than \numt{}.''
However, \msb is a model-choice estimator, and hence the goal is to estimate the
posterior probability of divergence models.
\citet{Oaks2012} demonstrate that even when all assumptions of the model are
met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
4), whereas \numt{} performs better.
Furthermore, \vmratio{} is limited to estimating the probability of only a
single divergence model (the one divergence model), and thus its utility for
model choice is extremely limited.

The model-choice utility of \vmratio{} is limited to the probability that
this continuous statistic, which can range from zero to infinity, is at its lower
limit of zero. In theory, this point density will always be zero, thus an
arbitrary threshold (0.01 is used throughout the \msb literature) must 
be chosen to make the probability estimable.
However, it is still not surprising to see that it is numerically difficult to
obtain reliable estimates of the probability that the continuous \vmratio{}
statistic is ``near'' its limit of zero.
It is much easier, less subjective, and more interpretable to estimate
the probability of the discrete parameter of the model, \numt{}, is
at its lower limit of one.
Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
estimator of model probability than \vmratio{}.

>>>>

Recently, the journal \emph{Evolution} has served as a medium for a discussion
of potential promises and pitfalls of approximate Bayesian methods of
comparative phylogeographical model choice and prior selection.
The discussion has centered around the model-choice method, \msb
\citep{Huang2011}, which uses approximate Bayesian computation (ABC) to
estimate the temporal distribution of divergences among co-distributed pairs of
taxa by sampling over models of divergence that range from a single
divergence-time parameter (i.e., all pairs of taxa diverged at the same time)
to the fully generalized model in which each pair is assigned a divergence-time
parameter.
A full description of the model is given in \citet{Oaks2012}.

\citet{Oaks2012} published results of simulation-based power analyses showing
that \msb is biased toward inferring models of temporally clustered divergence
times across taxon pairs.
\citet{Hickerson2013} published a response to \citet{Oaks2012} where they
present an empirical Bayesian model-averaging approach that they conclude
circumvents the problems revealed by \citet{Oaks2012}.
These papers are largely in agreement, with \citet{Hickerson2013} reiterating
much of the discussion of \citet{Oaks2012} regarding the impact of broad
uniform priors on Bayesian model choice.
Both papers agree that when broad uniform priors are used to reflect \emph{a
prior} uncertainty about the divergence times of all pairs of taxa under study,
the method can be biased toward inferring biogeographically interesting
scenarios of temporally clustered divergences among taxa (i.e. models with few
divergence-time parameters).
Both papers also agree that this bias toward models with few divergence-time
parameters is due to the much greater parameter space of models with more
divergence-time parameters.
However, the papers disgree about the primary mechanism by which the greater
parameter space of parameter-rich models causes the bias toward clustered
divergence models;
\citet{Oaks2012} suggest a statistical issue is the cause, whereas
\citep{Hickerson2013} prefer a numerical explanation.
% The main areas of disagreement between the papers are centered around
% (1) the mechanism by which broad uniform priors cause the bias of \msb toward
% models with few divergence-time parameters, and
% (2) how to potentially mitigate this issue.

\citet{Oaks2012} suggest the primary cause of the poor behavior of the method
is likely the low marginal likelihoods of parameter-rich models integrated over
vast parameter space with low probability of producing the data, yet relatively
high prior density \citep{Lindley1957}.
Note, this suggests the bias is extrinsic to \msb, and the numerical
approximation machinery of \msb could be sound.
\citet{Hickerson2013} suggest the bias is intrinsic to \msb, concluding the
method's rejection algorithm is inefficient and will be biased
toward models with fewer divergence-time parameters, because the parameter
space of these models are more densely sampled relative to models with more
divergence-time parameters.
We agree that sampling error is present in all numerical Bayesian estimation
methods, however, we show here that it is not a major contributer to the
biases found by \citet{Oaks2012}.
We present results of additional analyses that, when combined with
the results of \citet{Oaks2012}, strongly suggest small marginal likelihoods of
models with more parameter space is a primary mechanism by which uniform priors
are causing the bias of \msb.

% We show that this argument is based on empirical Bayesian reasoning and
% questionable assumptions, and does not explain the bias of the method found in
% many of the analyses of \citet{Oaks2012}.

There is also some disagreement between the papers also regarding potential
solutions to the bias.
\citet{Oaks2012} suggest alternative prior probability distributions on
divergence-times and other nuisance parameters could increase the marginal
likelihoods of models with more divergence-time parameters and thus reduce the
bias.
Alternatively, \citet{Hickerson2013} present an approach that uses empirically
informed priors and Bayesian model-averaging in an attempt to accommodate
uncertainty in selecting priors.
\citet{Hickerson2013} champion this strategy as a means of avoiding the
pitfalls raised by \citet{Oaks2012}.
In general, we agree with the use of Bayesian model averaging to obtain a
posterior estimate marginalized over uncertainty in prior choice.
However, we question whether adding an additional dimension of model choice
that samples over empirically informed uniform priors is a fruitful solution
for an already biased model-choice method.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different time units, which makes the results presented
in their response difficult to interpret (see Supporting Information for
details).
However, here we show that even when that error is avoided, the empirical
Bayesian model-averaging approach of \citet{Hickerson2013} remains biased
toward clustered divergence models, and is predisposed to excluding true
parameter space.
Their approach provides another means by which the model can ``escape'' large
parameter space, and the method's bias toward smaller models remains.
Importantly, this bias can manifest by preferring narrow, empirically informed
priors that exclude true parameter space.
Furthermore we discuss the potential theoretical and practical problems of
empirical Bayesian model choice.

% We discuss errors in this approach that render some of their results invalid
% and leave the remaining results difficult to interpret.
% Furthermore, we follow the recommendations of \citet{Oaks2012} and present
% simulation-based assessments of Hickerson et al.'s (\citeyear{Hickerson2013})
% approach.

@@

\section*{The potential implications of empirical Bayesian model choice}
\citet{Hickerson2013} repeatedly refer to the prior distributions used by
\citet{Oaks2012} as ``poorly selected.''
However, this is misleading, as \citet{Oaks2012} discuss in detail how they
selected their priors to reflect the large amount of uncertainty about the
model's parameters for their study system (see their section ``Specifying and
simulating the joint prior'').
\citet{Oaks2012} further discuss how the use of uniform prior distributions in
\msb for many of the model's parameters requires investigators to select broad
priors to avoid excluding the truth \textit{a priori}, and how these broad
uniform priors place too much density in improbable regions of parameter space.
What \citet{Hickerson2013} mean by ``poorly selected'' is that
\citet{Oaks2012} did not use their data to inform their initial prior
distributions (i.e., they did not use an empirical Bayesian approach).
\citet{Oaks2012} did use empirically informed priors when assessing both the
prior sensitivity of the method and the power of \msb under real-world
situations when a large amount of prior information is available. 
In doing so, \citet{Oaks2012} discuss
the potential dangers of taking an empirical Bayesian approach to model choice
(see the last paragraph of ``Assessing prior sensitivity of \msb'' in
\citet{Oaks2012}), and we expand on this here by exploring both the theoretical
and practical implications of using empirically informed priors for Bayesian
model choice.

@@

Whereas empirical Bayes approaches can provide powerful methods for parameter
estimation, a theoretical justification for empirical Bayes approaches to model
choice is questionable.
In Bayesian model choice, the primary goal is not to estimate parameters, but
to estimate the relative probabilities of candidate models.
In a simple example of Bayesian model choice where two candidate models, $M_1$
and $M_2$, are being compared, the goal is to estimate the posterior
probabilities of these two models.
Again, we can use Bayes' rule to calculate this as
\begin{equation}
    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
    M_2)p(M_2)}.
    \label{eq:bayesmodelchoice}
\end{equation}
By comparing Equations \ref{eq:bayesrule} and \ref{eq:bayesmodelchoice}, we
can see the fundamental differences between Bayesian parameter estimation and
model choice.
In Equation \ref{eq:bayesrule}, we can see that the posterior density of any
state $\myTheta{}$ of the model, is the prior density updated by the
probabiliity of the data given the state $\myTheta{}$ (the likelihood of
$\myTheta{}$).
The integral over the entire parameter space of the model, which is defined by
the priors, only appears as a normalizing constant in the denominator.
Thus, as long as the joint prior distribution contains the values of $\myTheta{}$
that maximize the probability of the data (i.e., the maximum likelihood) and the
data are strongly informative relative to the prior, the modes of the parameter
estimates will be relatively robust to prior choice, even if the shape of the
posterior is incorrect due to using the data to inform the priors.
This is why empirical Bayes can work well for estimating the values of model
parameters.

However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model
the updates the prior probability of that model to yield posterior
probability.
Thus the integral over the entire parameter space of the likelihood weighted by
the prior probability density is no longer a normalizing constant, but rather
is what informs the posterior probability of that model from the data.
As a result, Bayesian model choice tends to be much more sensitive to priors
than Bayesian parameter estimation.
Another important difference of Bayesian model choice that is illustrated by
Equations~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a parameter of the model, but rather a
single value that is a summary of the entire posterior distribution.
This further complicates the meaning of model posterior probabilities when
Bayes' rule is violated by empirically informing model priors.
Model posterior probabilities are inherently measures of our belief in
the models after our prior beliefs are updated by the data being analyzed.
Unlike model parameters, there are no true values for model probabilities,
and there are no measures of uncertainty in the posterior estimates. 
For empirical Bayesian parameter estimation, the hope is that by giving the
data more weight relative to the prior, the posterior distribution will be
peaked near the true value(s) of the model's parameter(s).
There is no such justification for empirical Bayesian model choice, because
there are no true model probabilities that we are trying to estimate.
Rather using the data twice will simply mislead our posterior
beliefs in the models being compared.

@@

### 2013-11-19 manuscripts/reply/reply-si-main.tex
Working on reply.
Moving error section to SI, and working on Intro.

>>>>

\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
The main parameter of the \msb model is the vector of divergence
times for each of the taxon pairs,
$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
\citep{Oaks2012}.
\citet{Hickerson2013} argue that the dispersion index of this vector,
\vmratio{}, is a better model-choice estimator than the number of 
divergence time parameters within the vector,
\numt{}.
They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
This plot shows, not surprisingly, that \vmratio{} has very little information
about the number of divergences among taxa.
Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
biogeographically relevant than \numt{}.
% We struggle to follow this logic.
Certainly the maximum information is contained within the divergence time
vector that \vmratio{} is summarizing.
Clearly, the number of divergence time parameters within the vector and their
values is more informative than its variance (i.e., the dispersion index is not
a sufficient statistic for \divtvector).

\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
better than \numt{}.''
However, \msb is a model-choice estimator, and hence the goal is to estimate the
posterior probability of divergence models.
\citet{Oaks2012} demonstrate that even when all assumptions of the model are
met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
4), whereas \numt{} performs better.
Furthermore, \vmratio{} is limited to estimating the probability of only a
single divergence model (the one divergence model), and thus its utility for
model choice is extremely limited.

The model-choice utility of \vmratio{} is limited to the probability that
this continuous statistic, which can range from zero to infinity, is at its lower
limit of zero. In theory, this point density will always be zero, thus an
arbitrary threshold (0.01 is used throughout the \msb literature) must 
be chosen to make the probability estimable.
However, it is still not surprising to see that it is numerically difficult to
obtain reliable estimates of the probability that the continuous \vmratio{}
statistic is ``near'' its limit of zero.
It is much easier, less subjective, and more interpretable to estimate
the probability of the discrete parameter of the model, \numt{}, is
at its lower limit of one.
Thus, it is not surprising that \citet{Oaks2012} find that \numt{} is a better
estimator of model probability than \vmratio{}.
